Sampling requirements to detect ecosystem change

### 2020-06-23 R/manuscript_draft.Rmd
updated intro paragraph

pacman::p_load(patchwork, stickylabeller, ggplot2, dplyr, wesanderson, reshape, cowplot)
```

@@

@@

Two dimensions are to be considered when sampling: spatial and temporal dimension. To estimate a same parameter, population abundance, optimal spatial sampling strategies are based on spatially balanced sampling [@kermorvant_optimizing_2019], while optimal temporal sampling strategies are more cost efficient with a targeted sampling, e.g., around the period of reproduction [@jackson_sampling_2008]. _Add at least another sentence here to make the transition to our work, but right now I can't think of a good formulation_

Here, we develop a set of tools to determine the appropriate number of samples and sampling approach when dealing with data sources where only a subset of samples are analyzed. We tailor our analysis to the detection of a changepoint in a time series, but our approach is applicable to other questions as well. Changepoints are an important characteristic of a time series as they can indicate a underlying change in ecosystem processes [@James2014]. We focus on paleoecological core samples as one example of this type of data. We examine the situation where the goal is to detect the time at which a significant change in an ecological community occurs, i.e. a changepoint. However, our approaches are widely applicable to other questions and data types. We first investigate these tools using a simulation-based approach. We then test the tools on a case study from a paleosequence of Cladocera community assemblage from Lake Varese located in the subalpine region of north-western Italy [@Bruel2018].

>>>>

pacman::p_load(patchwork, stickylabeller, ggplot2, dplyr, wesanderson, reshape, cowplot, vegan, ecp)
```

@@

Because ecological systems are dynamic in both space and time, it is essential that sampling designs account for spatio-temporal dynamics (Williams et al. 2018). To estimate a same parameter, population abundance, optimal spatial sampling strategies are based on spatially balanced sampling [@kermorvant_optimizing_2019], while optimal temporal sampling strategies are more cost efficient with a targeted sampling, e.g., around the period of reproduction [@jackson_sampling_2008]. In either case, sampling time and lcoations can be chosen in an iterative process to be cost-effective and reduce the uncertainty in the process (Hooten et al. 2009). With the ability to choose precisely when to sample, we can move beyond random, interval, or opportunistic sampling designs (citations). This is particuraly relevant in situations where a subset of samples already collected need to be analyzed.

@@

Building on ideas from dynamic survey design (citations), we develop a set of tools to determine the appropriate number of samples and sampling approach when dealing with data sources where only a subset of samples are analyzed. We tailor our analysis to the detection of a changepoint in a time series, but our approach is applicable to other questions as well. Changepoints are an important characteristic of a time series as they can indicate a underlying change in ecosystem processes [@James2014]. We focus on paleoecological core samples as one example of this type of data. We examine the situation where the goal is to detect the time at which a significant change in an ecological community occurs, i.e. a changepoint. However, our approaches are widely applicable to other questions and data types. We first investigate these tools using a simulation-based approach. We then test the tools on a case study from a paleosequence of Cladocera community assemblage from Lake Varese located in the subalpine region of north-western Italy [@Bruel2018].
